An introduction to Docker for reproducible research

### 2014-11-11 docker-paper/manuscript.Rmd
docker for travis, maybe?

Docker is an open source project that builds on many long-familiar
technologies from operating systems research: LXC containers,
virtualization of the OS, and a hash-based or git-like versioning
and differencing system, among others (see [docs.docker.com/faq](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc) for
an excellent overview of what Docker adds to plain LXC).

@@

Rather than put the burden on the researcher to adopt
a very different workflow, the researcher can thus use their familiar editors,
etc, while immediately benefiting from the fact their computations can
be easily deployed on the cloud or the machines of collaborators without
any further effort than installing Docker software. The docker approach
is particularly well suited for moving between local and cloud platforms
when a web-based integrated development environment is available, such as
RStudio Server, or (to lesser extent) an iPython notebook.

@@

First, Docker facilitates modular reuse by building one container on top
of another through the use of `FROM` directive in Dockerfiles. This acts
like a software dependency; but unlike other software, a Dockerfile must
have exactly one dependency (one `FROM` line). Particular version of
the dependency can be specified using the `:` notation, or ommitted to
default to the latest version. To install software, Dockerfiles leverage
existing package managers on common Linux distributions such as `apt`
on Debian. These also permit installing either specific or only the
latest version.

@@

```bash
docker run -d --name db training/postgres
docker run -d -P --link db:db training/webapp python app.py
```
This separates the task fo running the database (first line) from
the application used to analyze the data (second line), allowing a more
modular approach to reuse: another researcher could use the same database
container while connecting it to different scripts.

>>>>

Docker is an open source project that builds on many
long-familiar technologies from operating systems research:
LXC containers, virtualization of the OS, and a hash-based
or git-like versioning and differencing system, among others (see
[docs.docker.com/faq](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)
for an excellent overview of what Docker adds to plain LXC). The
official documentation [docs.docker.com](https://docs.docker.com)
already provides a thorough introduction in how to use Docker
software; here my focus is on describing the implications this
has for reproducible research.  Readers may also find it helpful
to see more detailed examples of using Docker to capture, share,
and interact with a specific computational environments. Some such
examples, along with more detailed documentation on use can be found at
[github.com/rocker-org](https://github.com/rocker-org/rocker).

@@

Rather than put the burden on the researcher to adopt
a very different workflow, the researcher can thus use their familiar editors,
etc., while immediately benefiting from the fact their computations can
be easily deployed on the cloud or the machines of collaborators without
any further effort than installing Docker software. The docker approach
is particularly well suited for moving between local and cloud platforms
when a web-based integrated development environment is available, such as
RStudio Server, or (to lesser extent) an iPython notebook.

@@

First, Docker facilitates modular reuse by building one container on top
of another through the use of `FROM` directive in Dockerfiles. This acts
like a software dependency; but unlike other software, a Dockerfile must
have exactly one dependency (one `FROM` line). Particular version of
the dependency can be specified using the `:` notation, or omitted to
default to the latest version. To install software, Dockerfiles leverage
existing package managers on common Linux distributions such as `apt`
on Debian. These also permit installing either specific or only the
latest version.

@@

```bash
docker run -d --name db training/postgres
docker run -d -P --link db:db training/webapp python app.py
```
This separates the task for running the database (first line) from
the application used to analyze the data (second line), allowing a more
modular approach to reuse: another researcher could use the same database
container while connecting it to different scripts.
